Some thought experiments on the applicability of Granger causality and Directed Information in inferring direction of information flows

### 2015-07-05 paper.tex
Add main points for Allerton draft.

>>>>

\subsection{Motivation}

\begin{enumerate}
	\item Increased interest in understanding (a) functional connectivity maps and (b) computational networks of the brain.
	\item Evidenced by funding - Brain Initiative, Human Brain project.
	\item Quote one of the missions of the Brain Initiative - esp. for connectivity maps, but also for computation
	\item Quote papers that try to understand how *information flows* in the brain. (past references from the ISIT paper - look for some more).
	\item Last line: "The question we raise is - Do these directed connectivity maps, as determined by directional causal influence measures such as Granger causality, correctly identify the directions along which information flows in the brain?"
\end{enumerate}

\subsection{How Granger Causality is used in Neuroscience today}

\begin{enumerate}
	\item We take Brovelli et.\ al.\ and Ding et.\ al.\ as the basis / canonical reference, as this appears to be generally accepted by the community (Cite review papers that cite these two papers; cite Anil Seth's MVGC package paper).
	\item Explain Granger causality
	\item Explain the exact method used - including trials, pre-processing steps, etc.\ - as described in the Bressler paper.
	\item Granger's original analysis does not compare GC on forward and reverse links, however, this appears to be the commonly done in practice (cite Brovelli and review papers if possible)
	\item State our problem statement again - we take issue not with any of these steps, but with the interpretation that if greater Granger causal influences are seen to be from X to Y, then information must be flowing from X to Y in the brain.
\end{enumerate}

\subsection{Our counter-examples and main result}

\begin{enumerate}
	\item Define Granger causal influence and directed info in English again. Define information flow (use examples, or define formally? Formal definition should come in the math sections). Describe the problem again - is the direction of causal influence the same as the direction of information flow?
	\item Describe the counter-examples in English.
		\begin{enumerate}
			\item Transmitter has the message (msg is known at the transmitter before it is known at the receiver - this is what distinguishes Tx from Rx). Therefore, the true direction in which information is flowing is *known* to us, by design of the experiment. This is the *ground truth*, against which we can compare the results of Granger causality.
			\item Feedback link is present - we consider two cases: feedback link noiseless, and feedback link noisy.
			\item We consider the case where Tx and Rx are using the S\&K strategy to communicate.
			\item We record time series of transmissions at Tx and Rx, and using this we use Granger Causality to try to understand which direction the information is flowing.
		\end{enumerate}
	\item We discover that for reasonable choices of parameter values, the direction of Granger causal influence is opposite to the direction of information flow.
\end{enumerate}

\subsection{Survey of previous criticisms of Granger Causality}

\begin{enumerate}
	\item Granger causality vs. true causality - cite Pearl. Our argument is, in spirit, very similar to the common criticism of Granger causality, which states that GC is a statistical quantity. It does not answer the question of true causal influence: if X is turned off, will Y behave differenty? (Need to quote from Pearl here).However, in the situation we are studying, we ask a different question - we wish to understand whether or not GC gives the correct direction of *information flow*. In our analysis, causal influences are present in both directions.
	\item Hidden node problem - hard to estimate GC correctly unless you are measuring all the nodes. (measurement problem)
	\item Noisy measurements can change the direction - cite Friston, cite that economics paper
\end{enumerate}

Note: The technical objections presented above (apart from GC-is-not-true-causality) indicate that estimating GC is hard, because it is hard to measure all nodes, etc. There is some deficiency in the measurement process. If we could measure better (by sampling more nodes, or with lesser noise, or...) then we could get the correct answer. Our objection, on the other hand, is much more fundamental. *Even if* AR coeffs are computed correctly, and the GC metric is perfectly computed (as is the case in our counter-examples), GC would *still* not give the correct direction of information flow. To our knowledge, this is a completely novel argument against GC, and one that is far more severe than previous objections, at least as far as the usage of GC in neuroscience is concerned.

Other objections we could add - Kun Zhang's paper (subsampling can cause errors), effective connectivity or plumbing? (venous drainage),

\subsection{Possible criticisms of this work}
\begin{enumerate}
	\item Circularity - there is no circularity of argument. We are just assuming that there are sufficient number of trials to fully estimate the coefficients of the non-stationary AR process. We describe how this might be achieved in section ... (maybe this should come at the end?)
	\item Math is not complicated - this is a strength of our analysis. Even simple counter-examples that do not employ difficult proof techniques are able to show that GC does not always correctly estimate the direction of info flow.
\end{enumerate}

\subsection{Shortcomings in our analysis}
\begin{enumerate}
	\item Not evolving source - not a stationary AR process. To circumvent this, we discuss estimation of non-stationary AR coeffs, since this is a neuroscience experiment, so we have many realizations of the same AR process in the form of multiple trials (cite Ding again. Quote if necessary) A future work will address this shortcoming by examining how GC works with an evolving source that produces a stationary process.
	\item We restrict ourselves to Gaussian noise for simplicity of analysis. Noise in the actual brain is likely to be very different. However, AR processes are often used to model ERPs (need a citation for this).
	\item In the S\&K algo, we have simplified power and energy constraints
	\item This is for a simple two-node network. But for larger networks, the issue of incorrectly estimating the direction will probably only get exacerbated.
\end{enumerate}

\subsection{Discussion}
\begin{enumerate}
	\item Some of this misconception might have arisen because of Massey's directed info conservation - we expect that dir info in the forward dirn goes up if the other one goes down - but this is a misnomer. (Do we still need this? I am increasingly of the opinion that directed info / GC capture nothing about the message)
\end{enumerate}

\section{ghgh}
